---
title: "2.1 Random Variables and Probability Distributions"
output: html_notebook
toc: yes
---

Generate random sample.

```{r}
sample <- sample(1:6, 1)
```

# Probability Distributions of Discrete Random Variables

Generate vector of probabilities. Set up the probability vector using the function `rep()`, which replicates a given value a specified number of times.

```{r}
probability <- rep(1/6, 6)

# Plot the probability
plot(
  probability,
  xlab = "Outcomes",
  ylab = "Probability",
  main = "Probability Distribution"
)
```

For the **cumulative probability distribution** we need the cumulative probabilities, i.e., we need the cumulative sums of the vector probability. These sums can be computed using `cumsum()`.

```{r}
cum_probability <- cumsum(probability)

plot(cum_probability,
     xlab = "Outcome",
     ylab = "Cumulative Probability",
     main = "Cumulative Probability Distribution")
```

# Bernoulli Trials

The set of elements from which `sample()` draws outcomes does not have to consist of numbers only. We might as well simulate coin tossing with outcomes\
$H$(heads) and $T$(tails).

```{r}
sample(c("H", "T"), 1)
```

The result of a single coin toss is a *Bernoulli* distributed random variable.

Example: toss a coin\
10 times in a row and wonder how likely it is to end up with a 5 times heads. This is a typical example of what we call a *Bernoulli experiment* as it consists of $n = 10$ Bernoulli trials that are independent of each other and we are interested in the likelihood of observing\
$k = 5$ successes $H$ that occur with probability $p = 0.5$ (assuming a fair coin) in each trial. Note that the order of the outcomes does not matter here.

It is a well known result that the number of successes $k$ in a Bernoulli experiment follows a binomial distribution. We denote this as

$$
k \sim B(n,p)
$$ The probability of observing $k$ success in the experiment $B(n, p)$ is given by $$
f(k)=P(k)=\left(\begin{array}{l}
n \\
k
\end{array}\right) \cdot p^k \cdot(1-p)^{n-k}=\frac{n !}{k !(n-k) !} \cdot p^k \cdot(1-p)^{n-k}
$$ with $\left(\begin{array}{l} n \\ k \end{array}\right)$ as the binomial coefficient.

In R, we use `dbinom()` which calculates `P(k|n,p)`, the probability of the binomial distribution given the parameters $x(k)$, size($n$) and prob($p$). Now compute $P(k=5|n=10, p=0.5)$, short handed as $P(k=5)$.

```{r}
dbinom(x = 5,
       size = 10,
       prob = 0.5
       )
```

Now calculate $P(4 \leq k \leq 7)$

```{r}
sum(dbinom(x = 4:7, size = 10, prob = 0.5))
```

An alternative approach is to use `pbinom()`, the distribution function of the binomial distribution to compute $P(4 \leq k \leq 7) = P(k \leq 7) - P(k \leq 3)$

```{r}
# compute P(4 <= k <= 7) using 'pbinom()'
pbinom(size = 10, prob = 0.5, q = 7) - pbinom(size = 10, prob = 0.5, q = 3)
```
The probability distribution of a discrete random variable is nothing but a list of all possible outcomes that can occur and their respective probabilities. In the coin tossing example we have  
11 possible outcomes for $k$.

```{r}
k <- 0:10
k
```
To visualise.
```{r}
probability <- dbinom(x = k,
                      size = 10,
                      prob = 0.5)

# Plot the outcome
plot(x = k,
     y = probability,
     main = "Probability Distribution Function")
```
Also plot the cumulative distribution function of $k$.
```{r}
prob <- pbinom(q = k,
               size = 10,
               prob = 0.5)

# Plot
plot(x = k,
     y = prob,
     main = "Cumulative Distribution Function")
```

# Expected Value, Mean and Variance
